\section{Character Tables}

In Example~\ref{Ch2:Eg:Cycl_3_Char}, we constructed a table consisting of the character values of each conjugacy class of the cyclic group of order $3$ (in this case, each class consisted of only one element---but this is not always the case). Our approach in Example~\ref{Ch2:Eg:Cycl_3_Char} was direct computation, which, while straightforward, is not always practical---especially when dealing larger groups. In this section, we will introduce a more systematic way of constructing such tables using the orthogonality properties of irreducible characters.

\begin{boxnotation}
    For the purposes of this section, let $C_1, \ldots, C_k$ be the conjugacy classes of $G$ with representatives $g_1, \ldots, g_k$ respectively. Recall that by the Orbit-Stabiliser Theorem,
    \begin{align*}
        \abs{C_i} &= \frac{\abs{G}}{\abs{C_G(g_i)}}
    \end{align*}
    where $C_G(\cdot)$ refers to the centraliser\footnote{Recall that the centraliser of a group element is the set of all elements that commute with it.} of an element in $G$. Now, let $\Irr{G} = \set{\chi_1, \ldots, \chi_k}$.
    \end{boxnotation}

\subsection{The Orthogonality Relations}

\begin{boxdefinition}[Character Table]
    A character table is a $k \times k$ table whose columns are the conjugacy classes (or their representatives) and whose rows are the irreducible characters.
    \begin{table}[H]
        \centering
        \begin{tabular}{c|ccc}
            & $g_1$ & $\cdots$ & $g_k$ \\
            \hline
            $\chi_1$ & $\chiof{1}{g_1}$ & $\cdots$ & $\chiof{1}{g_k}$ \\
            $\vdots$ & $\vdots$ & $\ddots$ & $\vdots$ \\
            $\chi_k$ & $\chiof{k}{g_1}$ & $\cdots$ & $\chiof{k}{g_k}$
        \end{tabular}
    \end{table}
    In other words, it is a table whose $\parenth{i,j}$th entry is $\chi_i(g_j)$.
    % Maybe draw this...
\end{boxdefinition}

The table in Example~\ref{Ch2:Eg:Cycl_3_Char} is an example of a character table.

It turns out that we have a certain orthogonality property as we fix two rows and sum across the columns of a character table.

\begin{boxproposition}[First Orthogonality Relation]\label{SP:1st_Orth_Rel}
    Given $1 \leq r, s \leq k$, we have
    \begin{align*}
        \frac{1}{\abs{G}} \sum_{j=1}^{k} \abs{C_j} \chi_r\!\parenth{g_j} \chi_s\!\parenth{g_j\inv} &= \delta_{rs}
    \end{align*}
\end{boxproposition}
\begin{proof}
    The proof is quite straightforward. It merely involves applying the centrality of characters to the definition of the inner-product and concluding using the Orthogonality Theorem:
    \begin{align*}
        \delta_{rs} &=
        \cycl{\chi_r, \chi_s} \\
        &= \frac{1}{\abs{G}} \sum_{g \in G} \chi_r(g) \chi_s\!\parenth{g\inv} \\
        &= \frac{1}{\abs{G}} \sum_{j=1}^{k} \parenth{\sum_{g \in C_j} \chi_r(g) \chi_s\!\parenth{g\inv}} \\
        &= \frac{1}{\abs{G}} \sum_{j=1}^{k} \abs{C_j} \chi_r\!\parenth{g_j} \chi_s\!\parenth{g_j\inv}
    \end{align*}
    where the last inequality follows from the fact that characters are central (ie, they are equal for all elements of a conjugacy class).
\end{proof}
\begin{remark}
    Note that the result of Proposition~\ref{SP:1st_Orth_Rel} can be equivalently phrased in the following manner: given $1 \leq r, s \leq k$,
    \begin{align}
        \sum_{j=1}^{k} \frac{1}{\abs{C_G(g_j)}} \chi_r\!\parenth{g_j} \chi_s\!\parenth{g_j\inv} &= \delta_{rs}
        \label{SP:Eq:1st_Orth_Rel_Conj_Cl}
    \end{align}
\end{remark}

\begin{comment}
We can use this to prove an important result about the regular representation.

\begin{theorem}\label{Ch2:Thm:Reg_Rep_Col_1}
    Let $\chi_{\reg}$ denote the character of the regular representation of $G$ over $\C$. Then,
    \begin{align*}
        \chi_{\reg} &= \sum_{j=1}^{k} \chiof{j}{1} \chi_j
    \end{align*}
\end{theorem}
\begin{proof}
    
    \verb|sorry|  % From phone
\end{proof}

\begin{corollary} \label{Ch2:Cor:Reg_Rep_Col_1_Grp_Size}
    $\sum_{i=1}^{k} \parenth{\chiof{i}{1}}^2 = \abs{G}$.
\end{corollary}
\begin{proof}
    We evaluate both sides of the expression in Theorem~\ref{Ch2:Thm:Reg_Rep_Col_1} at $1$ to obtain that
    \begin{align*}
        \chiof{\reg}{1} = \sum_{i=1}^{k} \parenth{\chi_i(1)}^2 = \abs{G}        
    \end{align*}
\end{proof}
\end{comment}

Indeed, we have a similar result about fixing two columns and summing down the rows.

\begin{boxproposition}[Second Orthogonality Relation]\label{SP:2nd_Orth_Rel}
    Given $1 \leq r, s \leq k$, we have
    \begin{align*}
        \sum_{i=1}^{k} \chiof{i}{g_r} \chiof{i}{g_s\inv} =
        \begin{cases}
            0 &\text{ if } r \neq s \\
            \abs{C_G(g_r)} &\text{ if } r = s
        \end{cases}
    \end{align*}
\end{boxproposition}
\begin{proof}
    Let $A$ be the matrix $\parenth{A_{ij}}_{1 \leq i, j \leq k}$, where $A_{ij} := \chiof{i}{g_j}$. Similarly, let $B$ be the matrix $\parenth{B_{ij}}_{1 \leq i, j \leq k}$, where $B_{ij} := \frac{\abs{C_i}}{\abs{G}} \chiof{j}{g_i\inv}$. Then, writing $AB = \parenth{\parenth{AB}_{pq}}_{1 \leq p, q \leq k}$, we have
    \begin{align*}
        \parenth{AB}_{pq} &=
        \sum_{l=1}^{k} A_{pl} B_{lq} \\
        &= \frac{1}{\abs{G}} \sum_{l=1}^{k} \abs{C_l} \chiof{p}{g_l} \chiof{q}{g_l\inv} \\
        &= \frac{1}{\abs{G}} \sum_{g \in G} \chi_p(g) \chi_q(g\inv) \\
        &= \cycl{\chi_p, \chi_q} = \delta_{pq}
    \end{align*}
    with the third inequality holding because all characters are class functions, meaning that terms in summation on the third line such that $g$ is conjugate to $g_l$ will be repeated $\abs{C_l}$ times.
    
    The above computation proves that $AB = I$, the identity matrix. This, in particular, means that $BA = I$ as well. In particular, writing $BA = \parenth{\parenth{BA}_{pq}}_{1 \leq p, q \leq k}$, we see that
    \begin{align}
        \delta_{pq} = \parenth{BA}_{pq} &= \sum_{l=1}^{k} B_{pl} A_{lq} \nonumber \\
        &= \sum_{l=1}^{k} \frac{\abs{C_p}}{\abs{G}} \chiof{l}{g_p\inv} \chiof{l}{g_q} \nonumber \\
        &= \sum_{j=1}^{k} \frac{1}{\abs{C_G(g_r)}} \chiof{i}{g_r} \chiof{i}{g_s\inv}
        \label{SP:Eq:2nd_Orth_Rel_Conj_Cl}
    \end{align}
    A simple rearrangement then yields the desired result.
\end{proof}
\begin{remark}
    Observe that \eqref{SP:Eq:2nd_Orth_Rel_Conj_Cl} bears a striking similarity to \eqref{SP:Eq:1st_Orth_Rel_Conj_Cl}. The only difference is that in \eqref{SP:Eq:1st_Orth_Rel_Conj_Cl}, we fix the character and sum over the conjugacy classes (ie, fix the row and sum across the columns), while in \eqref{SP:Eq:2nd_Orth_Rel_Conj_Cl}, we fix the conjugacy class and sum over the characters (ie, fix the column and sum across the rows).
\end{remark}

In the next subsection, we will demonstrate how we can use these orthogonality relations, in combination with other results---most notably, \eqref{Ch2:Eq:Grp_Order_Sum_Char_Sq}---to construct character tables.

\subsection{Constructing Character Tables}

First, we construct the character table of $S_3$ using relatively unsophisticated machinery.

\begin{boxexample}[$S_3$]\label{Ch2:Eg:S3_CharTable}
    $S_3$ has precisely three conjugacy classes $C_1$, $C_2$ and $C_3$ with representatives $g_1 = 1$, $g_2 = \parenth{1 2}$ and $g_3 = \parenth{1 2 3}$. We then have the following character table for $S_3$:
    \begin{table}[H]
        \centering
        \begin{tabular}{c|ccc}
            & $1$ & $\parenth{12}$ & $\parenth{123}$ \\
            \hline
            $\chi_1$ & $1$ & $1$ & $1$ \\
            $\chi_2$ & $1$ & $-1$ & $1$ \\
            $\chi_3$ & $2$ & $0$ & $-1$
        \end{tabular}
    \end{table}
    Here, $\chi_1$ corresponds to the trivial representation (of degree $1$), $\chi_2$ to the sign representation (of degree $1$), and $\chi_3$ to the subrepresentation $W = \Span{e_1 - e_2, e_1 - e_3}$ (of degree $2$) of the permutation representation on $\C^3$. Note that all of these representations are irreducible, the first two because their degrees are $1$, and the third because we can show $\cycl{\chi_W, \chi_W}$ to be equal to $1$ (cf. the fourth point of Proposition \ref{Ch2:Prop:Bhv_Irred_Char}).
\end{boxexample}

The maxim of `the larger the object, the more involved the computations' holds no less true for character tables. To construct the character table of $S_4$, we will need to use two ingredients that we did not require to construct the character table of $S_3$. The first consists of the Orthogonality Relations, and the second of the following result on lifting characters of quotients, which we will be able to apply to lift characters of $S_3$ to characters of $S_4$.

\begin{proposition}[Lifting of Characters]\label{Ch2:Prop:Lifting_Chars}
    Let $N \nsg G$. If $\chi : \quotient{G}{N} \to \C$ is a character of $\quotient{G}{N}$, then the map $\chi' : G \to \C : g \mapsto \chi(gN)$ is a character of $G$. In particular, $\chi$ is irreducible if and only if $\chi'$ is irreducible.
\end{proposition}
\begin{proof}
    The result follows immediately from Proposition~\ref{Ch1:Prop:Lifting}.
\end{proof}

\begin{boxexample}[$S_4$]
    Observe that the conjugacy classes $C_i$, $i = 1, \ldots, 5$, of $S_4$ correspond precisely to the various possible cycle shapes. They therefore have representatives $g_1 = 1$, $g_2 = \parenth{12}$, $g_3 = \parenth{123}$, $g_4 = \parenth{12}\parenth{34}$, and $g_5 = \parenth{1234}$. The conjugacy classes have sizes $1$, $6$, $8$, $3$ and $6$ respectively. As there are five conjugacy classes, there are, in particular, exactly five irreducible characters of $S_4$. We can compute the values of four of them with relative ease, but for the fifth, we will, for now, fill in the table with placeholder variables:
    \begin{table}[H]
        \centering
        \begin{tabular}{c|ccccc}
            & $1$ & $\parenth{12}$ & $\parenth{123}$ & $(12)(34)$ & $(1234)$ \\
            \hline
            $\chi_1$ & $1$ & $1$ & $1$ & $1$ & $1$ \\
            $\chi_2$ & $1$ & $-1$ & $1$ & $1$ & $-1$ \\
            $\chi_3$ & $3$ & $1$ & $0$ & $-1$ & $-1$ \\
            $\chi_4$ & $2$ & $0$ & $-1$ & $2$ & $0$ \\
            $\chi_5$ & $a$ & $b$ & $c$ & $d$ & $e$
        \end{tabular}
    \end{table}
    We now describe each character and how the above values were computed. In the case of $\chi_5$, we will solve for $a$, $b$, $c$, $d$ and $e$ using the Orthogonality Relations.
    \begin{enumerate}
        \item $\chi_1$ is the \underline{trivial character}. It takes the value $1$ on all elements of $S_4$.
        
        \item $\chi_2$ is the \underline{sign character}\footnote{ie, that corresponding to the sign representation}. It takes the value $1$ on all even permutations and $-1$ on all odd permutations---indeed, it is exactly equal to the sign homomorphism.
        
        \item $\chi_3$ is the \underline{deleted permutation character} given by the subrepresentation of the permutation representation\footnote{ie, that corresponding to its action on $\C^4$ by permuting the standard basis} of $S_4$ corresponding to $\Span{e_1 - e_2, e_2 - e_3, e_3 - e_4}$, an $S_4$-invariant subspace of $\C^4$. The values it takes can be computed by taking the traces of the orthogonal matrices corresponding to the action of $S_4$ on this subspace.
        
        \item $\chi_4$ is given by \underline{lifting a character of $S_3$}. Consider
        \begin{align*}
            N = \set{1, (12)(34), (13)(24), (14)(23)} \nsg S_4
        \end{align*}
        One can show that $\quotient{S_4}{N} \cong S_3$. We can then apply Proposition~\ref{Ch2:Prop:Lifting_Chars}, taking advantage of the irreducibility equivalence, to lift the three irreducible characters from Example~\ref{Ch2:Eg:S3_CharTable} to irreducible characters of $S_4$. The lifts of the first two are $\chi_1$ and $\chi_2$, so we choose $\chi_4$ to be a lift of the third.
        
        \item $\chi_5$ is the \underline{regular character}. We can solve for $a,b,c,d,e$ using the Orthogonality Relations. From \eqref{Ch2:Eq:Grp_Order_Sum_Char_Sq}, we know that
        \begin{align*}
            1 + 1 + 9 + 4 + a^2 = 15 + a^2 = 24 = \abs{S_4}
        \end{align*}
        from which we can conclude that $a = \pm 3$. Seeing as $\chi_5(1)$ must correspond to the dimension of a representation, $a$ must be positive. So, $a = 3$. Then, applying the Second Orthonality Relation (Proposition~\ref{SP:2nd_Orth_Rel}) to the first two columns,
        \begin{align*}
            1 - 1 + 3 + 0 + ab = 3 + 3b = 0
        \end{align*}
        from which we can conclude that $b = -1$. In similar fashion, we can apply the Second Orthogonality Relation to the second and third columns to solve for $c$:
        \begin{align*}
            1 - 1 + 0 + 0 - c = 0 - c = 0
        \end{align*}
        giving $c = 0$. To solve for $d$, it does not make sense to apply the Second Orthogonality Relation to the third and fourth columns, because $cd = 0$, but we can certainly apply it to the \textit{second} and fourth columns:
        \begin{align*}
            1 - 1 - 1 + 0 - d = -1 - d = 0
        \end{align*}
        giving $d = -1$. Finally, we can apply the Second Orthogonality Relation to the fourth and fifth columns to solve for $e$:
        \begin{align*}
            1 - 1 + 1 + 0 - e = 1 - e = 0
        \end{align*}
        giving $e = 1$.
    \end{enumerate}
    Thus, the character table of $S_4$ is as follows:
    \begin{table}[H]
        \centering
        \begin{tabular}{c|ccccc}
            & $1$ & $\parenth{12}$ & $\parenth{123}$ & $(12)(34)$ & $(1234)$ \\
            \hline
            $\chi_1$ & $1$ & $1$ & $1$ & $1$ & $1$ \\
            $\chi_2$ & $1$ & $-1$ & $1$ & $1$ & $-1$ \\
            $\chi_3$ & $3$ & $1$ & $0$ & $-1$ & $-1$ \\
            $\chi_4$ & $2$ & $0$ & $-1$ & $2$ & $0$ \\
            $\chi_5$ & $3$ & $-1$ & $0$ & $-1$ & $1$
        \end{tabular}
    \end{table}
    We can double-check our computations for the regular character by applying the First Orthogonality Relation (Proposition~\ref{SP:2nd_Orth_Rel}) to the first and last rows (ie, by showing $\cycl{\chi_1, \chi_5} = 0$): clearly,
    \begin{align*}
        \frac{1}{\abs{G}} \sum_{i=1}^{5} \abs{C_i} \chiof{1}{g}\chiof{5}{g\inv} &= \frac{1}{24} \parenth{1 \cdot \bar{a} + 6 \cdot \bar{b} + 8 \cdot \bar{c} + 3 \cdot \bar{d} + 6 \cdot \bar{e}} \\
        &= \frac{1}{24}\parenth{3 - 6 + 0 - 3 + 6} = 0
    \end{align*}
    We do not prove here that the conjugacy classes have sizes $1$, $6$, $8$, $3$ and $6$ respectively, but this can be done by direct computation.
\end{boxexample}

Note that the number of irreducible characters of degree $1$ of any group $G$ is $\abs{G} / \abs{G'}$, where $G'$ is the derived subgroup of $G$. Indeed, there is a one-to-one correspondence between degree $1$ irreducible characters of $G$ and those of $\quotient{G}{G'}$.

It turns out that we can glean even more information about groups from their characters, for which we will need the properties of the algebraic integers.