\section{Integrality}

In this section, we exploit our understanding of the complex numbers to gain insight into the characters of finite groups.

\subsection{The Algebraic Integers}

First, we recall the notion of integrality over an integral domain.

\begin{boxdefinition}[Integrality]
    Let $R$ be an integral domain, and $S \supset R$ an extension. We say that $s \in S$ is integral over $R$ if one of the following equivalent conditions is satisfied:
    \begin{itemize}
        \item $s$ is a root of a monic polynomial in $R[X]$.
        \item The minimal polynomial of $s$ over $\Frac{R}$ is actually in $R[X]$.
    \end{itemize}
\end{boxdefinition}

We now define what it means for a complex number to be an algebraic integer.

\begin{boxdefinition}[Algebraic Integer]
    We say that a number $\alpha \in \C$ is an algebraic integer if it is integral over $\Z$.
\end{boxdefinition}

One can show that the set of all algebraic integers forms a ring. Indeed, this is nothing but the integral closure of $\Z$ in $\C$. This has an important consequence as far as character theory is concerned.

\begin{proposition}\label{Ch2:Prop:Char_Val_Alg_Int}
    For all characters $\chi$ of $G$ and $g \in G$, $\chi(g)$ is an algebraic integer.
\end{proposition}
\begin{proof}
    Fix $g \in G$ and a character $\chi$. Let $d = \ord{g}$. The fourth point of Proposition~\ref{Ch2:Prop:Bhv_Char} tells us that $\chi(g)$ is a sum of $d$th roots of unity, which, as roots of $X^d - 1 \in \Z[X]$, are certainly algebraic integers. Thus, as a sum of algebraic integers, $\chi(g)$ is an algebraic integer. 
\end{proof}

We now proceed to prove several results on the conjugacy classes of $G$. We begin by defining a simple way to obtain a central element of $\C G$ from a conjugacy class of $G$.

\begin{boxnotation}
    Given a conjugacy class $C$ of $G$, we define
    \begin{align}
        \hat{C} := \sum_{g \in C} g
        \label{Ch2:Eq:Hat_C}
    \end{align}
    as an element of $\C G$.
\end{boxnotation}

\begin{lemma}
    For any conjugacy class $C$ of $G$, $\hat{C} \in \Zof{\C G}$.
\end{lemma}
\begin{proof}
    It is clear that $\hat{C}$ commutes with every element of $\C \subset \C G$. We proceed to show that they also commute with elements of $G \subset \C G$. Let $g \in C$ be a representative of $C$. We know that every term in the sum in \eqref{Ch2:Eq:Hat_C} (ie, every element of $C$) is of the form $hgh\inv$ for some $h \in H$. For some $j \in G$, we have that
    \begin{align*}
        j\parenth{hgh\inv} = \parenth{\underbrace{j h g h\inv j\inv}_{\in C}} j
    \end{align*}
    Hence, left-multiplication by $j$ acts as a permutation of $C$. Therefore, $j \hat{C} = \hat{C} j$.
\end{proof}

\begin{lemma}\label{Ch2:Lem:Scalar_Action_Alg_Int}
    Let $g \in G$ and $C$ be the conjugacy class of $g$.
    Let $S$ be a simple \CGM\ with associated character $\chi_S$. The action of $\hat{C}$ on $S$ (as an element of $\C G$) can be simplified to one of scalar multiplication by a complex number $\lambda$:
    \begin{align*}
        \hat{C} \cdot s = \lambda s
    \end{align*}
    where $\displaystyle \lambda % = \frac{\abs{C}}{\abs{C_G(g)}} \frac{\chi(g)}{\chi_s(1)}
    = \abs{C} \frac{\chi_s(g)}{\chi_s(1)}$.
\end{lemma}
\begin{proof}
    Define a function $\phi : S \to S : s \mapsto \hat{C} \cdot s$. Since $\hat{C} \in \Zof{\C G}$, we have that $\forall x \in G$, $x \cdot \phi(s) = \phi(x \cdot s)$. This makes $\phi$ a \CGM\ homomorphism. Then, by Corollary~\ref{SP:Cor:Schur_fin_G_over_C_Morhp_Ver}, we know that $\phi = \lambda \id$, where $\lambda = \Tr{\phi} / \pdim{S}$. Note that $\pdim{S} = \chiof{S}{1}$. Further, if we write $\phi_h : S \to S : s \mapsto h \cdot s$ for $h \in C$, then we can express $\phi$ as $\sum_{h \in C} \phi_h$. Then,
    \begin{align*}
        \Tr{\phi} &= \sum_{h \in C} \Tr{\phi_h} \\
        &= \sum_{h \in G} \chiof{S}{h} \\
        &= \abs{C} \chiof{S}{g}
    \end{align*}
    where $g$ was our chosen representative of $C$. Therefore, $\displaystyle \lambda = \abs{C} \frac{\chi_s(g)}{\chi_s(1)}$.
\end{proof}

\begin{lemma}\label{Ch2:Lem:Alg_Int_Eigen}
    Let $G = \set{g_1, \ldots, g_n}$. Define $r := \sum_{i=1}^{n} \alpha_i g_i$ for some $\alpha_i \in \Z$. Suppose that $\exists \lambda, v \in \C G \setminus \set{0}$ such that $rv = \lambda v$. Then, $\lambda$ is an algebraic integer.
\end{lemma}
\begin{proof}
    For all $1 \leq j \leq n$,
    \begin{align*}
        r g_j = \sum_{i=1}^{n} \alpha_{i} g_i g_j = \sum_{i=1}^{n} \alpha_{ij} g_i
    \end{align*}
    where $\parenth{\alpha_{ij}}_{1 \leq i \leq n}$ is a permutation of the $\alpha_i$ induced by the action of $g_j$. This means, in particular, that $\alpha_{ij} \in \Z$ for all $1 \leq i,j \leq n$. Suppose $\exists \lambda, v \in \C G \setminus \set{0}$ such that $rv = \lambda v$. Writing $v = \sum_{j=1}^{n} \mu_i g_i$ for $\mu_i \in \C$, we can see that
    \begin{align*}
        rv = \sum_{j=1}^{n} \mu_j r g_j = \sum_{j=1}^{n} \mu_j \sum_{i=1}^{n} \alpha_{ij} g_i = \sum_{i=1}^{n} g_i \sum_{j=1}^{n} \alpha_{ij} \mu_j = \sum_{i=1}^{n} g_i \lambda \mu_i = \lambda v
    \end{align*}
    This tells us that $\lambda \mu_i = \sum_{j=1}^{n} \alpha_{ij} \mu_j$. In matrix form,
    \begin{align*}
        \begin{bmatrix}
            \alpha_{11} & \cdots & \alpha_{1n} \\
            \vdots & \ddots & \vdots \\
            \alpha_{n1} & \cdots & \alpha_{nn}
        \end{bmatrix}
        \begin{bmatrix}
            \mu_1 \\ \vdots \\ \mu_n
        \end{bmatrix}
        = \lambda
        \begin{bmatrix}
            \mu_1 \\ \vdots \\ \mu_n
        \end{bmatrix}
    \end{align*}
    This makes $\lambda$ is an eigenvalue of the matrix $A := \parenth{\alpha_{ij}}_{1 \leq i,j \leq n} \in \mat{n}{n}{\Z}$, and thereby, a (complex) root of the characteristic polynomial of $A$. This polynomial is monic, with coefficients in $\Z$, making $\lambda$ integral over $\Z$, and thereby, an algebraic integer.
\end{proof}

The point of the above Lemmas is the following Corollary.

\begin{corollary}\label{Ch2:Cor:lambda_Alg_Int}
    For any $\chi \in \Irr{G}$ and $g \in G$, the quantity $\lambda = \frac{\abs{G}}{\abs{C_G(g)}} \frac{\chi(g)}{\chi(1)} = \abs{C} \frac{\chi(g)}{\chi(1)}$ is an algebraic integer, where $C$ is the conjugacy class of $g$.
\end{corollary}
\begin{proof}
    % Recall that $C_G(g)$ is the centraliser of $g$. If $C$ is the conjugacy class of $g$, then $\frac{\abs{G}}{\abs{C_G(x)}} = \abs{C}$. Hence, $\lambda = \abs{C} \frac{\chi(g)}{\chi(1)}$.
    Lemma~\ref{Ch2:Lem:Scalar_Action_Alg_Int} tells us that $s \mapsto \lambda s$ is, indeed, a \CGM\ homomorphism. We can then apply Lemma~\ref{Ch2:Lem:Alg_Int_Eigen} to $\lambda$ and any $s \in S \setminus \set{0}$ to conclude that $\lambda$ is an algebraic integer.
\end{proof}

We can now show that the dimension of any irreducible representation of $G$ must divide $\abs{G}$.

\begin{boxtheorem}\label{Ch2:Thm:Char_div_Ord_Grp}
    For all $\chi \in \Irr{G}$, $\chi(1) \vert \abs{G}$.
\end{boxtheorem}
\begin{proof}
    It is a well-known fact that the integral closure of $\Z$ in $\Q$ is $\Z$ itself\footnote{ie, the only elements of $\Q$ that are integral over $\Z$ are the integers themselves.}. Hence, if we can show that $\frac{\abs{G}}{\chi(1)}$ is an algebraic integer, the fact that it is also rational will guarantee that it is an integer. We exploit the fact that $\cycl{\chi, \chi} = 1$ to write
    \begin{align*}
        \frac{\abs{G}}{\chi(1)} &= \frac{\abs{G}}{\chi(1)} \cycl{\chi, \chi} \\
        &= \frac{\cancel{\abs{G}}}{\chi(1)} \frac{1}{\cancel{\abs{G}}} \sum_{g \in G} \chi(g)\chi\!\parenth{g\inv}
    \end{align*}
    If $G$ has conjugacy classes $C_1, \ldots, C_r$ with representatives $g_1, \ldots, g_r$, then we can write
    \begin{align*}
        \frac{\abs{G}}{\chi(1)} &= \sum_{i=1}^{r} \frac{\abs{C_i}}{\chi(1)}\chi\!\parenth{g_i}\chi\!\parenth{g_i\inv}
    \end{align*}
    From Corollary~\ref{Ch2:Cor:lambda_Alg_Int} and Proposition~\ref{Ch2:Prop:Char_Val_Alg_Int}, we know that both $\frac{\abs{C_i}}{\chi(1)}\chi\!\parenth{g_i}$ and $\chi\!\parenth{g_i\inv}$ are algebraic integers. Hence, we can conclude that $\frac{\abs{G}}{\chi(1)}$ is an algebraic integer, and therefore, an integer.
\end{proof}

Corollary~\ref{Ch2:Cor:lambda_Alg_Int} also gives us another way to prove the specialisation of Proposition~\ref{Ch2:Prop:Char_Val_Alg_Int} to irreducible characters.

\begin{corollary}
    For any $\chi \in \Irr{G}$, $\chi(g)$ is an algebraic integer for all $g \in G$.
\end{corollary}
\begin{proof}
    Fix $g \in G$. We know that $\frac{\abs{G}}{\abs{C_G(g)}}\frac{\chi(g)}{\chi(1)}$ is an algebraic integer. Furthermore, we know that $\frac{\abs{G}}{\abs{C_G(g)} \cdot \chi(1)} \in \Q^{\times}$, and that the integral closure of $\Z$ in $\C$ (ie, the set of all algebraic integers) is a ring. Hence, $\chi(g) = \parenth{\frac{\abs{G}}{\abs{C_G(g)} \cdot \chi(1)}}\inv \frac{\abs{G}}{\abs{C_G(g)}}\frac{\chi(g)}{\chi(1)}$ is an algebraic integer.
\end{proof}



\subsection{The $n$th Roots of Unity}

We begin by stating a number theoretic result.

\begin{lemma}\label{Ch3:Lem:Int_Sum_Coprime_Roots_Unity}
    Let $w$ be an $n$th root of unity in $\C$. Then,
    \begin{align*}
        \sum_{\substack{1 \leq i \leq n \\ \parenth{i,n} = 1}} w^i \in \Z
    \end{align*}
\end{lemma}
\begin{proof}
    EXERCISE. Hint: induction on $n$.
\end{proof}

\begin{proposition}
    Let $g \in G$ be of order $n$. Suppose that $g^i$ is conjugate to $g$ for all $1 \leq i \leq n$ such that $\parenth{i,n} = 1$. Then, for all characters $\chi$ of $G$, $\chi(g)$ is an integer.
\end{proposition}
\begin{proof}
    Let $\chi$ be a character of $G$, with associated representation $\Vp$ of degree $m$ over $\C$. We know that $\chi(g) = \sum_{i=1}^{m} w_i$, where each $w_i$ is an $n$th root of unity (cf. the fourth point of Proposition~\ref{Ch2:Prop:Bhv_Char}). Indeed, the idempotent linear map $\rho(g)$ is diagonalisable, with matrix $\diag{w_1, \ldots, w_m}$ with respect to an appropriate basis (cf. Proposition~\ref{Ch1:Prop:Diag_Roots_of_Ord}). Hence, we have that $\rhoof{g^i} = \diag{w_1^i, \ldots, w_m^i}$, meaning that $\chi\!\parenth{g^i} = \sum_{j=1}^{m} w_j^i$. But this is nothing but $\sum_{1 \leq i \leq m, \parenth{i,n} = 1} w^i$, by assumption. Hence, $\chi(g)$ is an integer, by Lemma \ref{Ch3:Lem:Int_Sum_Coprime_Roots_Unity}.
\end{proof}

% \subsection{Congruence Conditions}

\begin{lemma} \label{Ch2:Lem:p_and_p'_parts}
    Fix $g \in G$, and let $p$ be a prime number. Then, $\exists! x, y \in G$ such that
    \begin{enumerate}[label= \normalfont \arabic*., noitemsep]
        \item $g = xy = yx$
        \item $\ord{x} = p^k$ for some $k \in \N$
        \item $\parenth{\ord{y}, p} = 1$
    \end{enumerate}
\end{lemma}
\begin{proof}
    Let $\ord{g} = u p^v$, where $\parenth{u,v = 1}$. By Bézout's Lemma, we know that $\exists a, b \in \Z$ such that $au + bp^v = 1$. % WHY?
    Then, we can define $x = g^{au}$ and $y = g^{bp^v}$. This satisfies all the conditions:
    \begin{enumerate}
        \item $g = xy = g^{au} g^{bp^v} = g^{bp^v} g^{au} = yx$.
        \item $x^{p^v} = 1$. % SO WHAT?
        \item $y^v = 1$.    % WHY?
    \end{enumerate}
    One can also show $x$ and $y$ to be the only elements of $G$ satisfying these conditions. %% SHOW!!!! 
\end{proof}

\begin{boxdefinition}[The $p$- $p'$-parts of $g$]
    For any $g \in G$, call $x$ and $y$ from Lemma \ref{Ch2:Lem:p_and_p'_parts} the $p$- and $p'$-parts of $g$ respectively.
\end{boxdefinition}

For the remainder of this subsection, let $n$ = $\abs{G}$, and let $\zeta = e^{\frac{2\pi i}{n}}$ be a primitive $n$th root of unity. Define $\Z[\zeta]$ to be the subring of $\C$ generated by $\Z$ and $\zeta$. Let $p$ be a prime number and $p\Z[\zeta]$ be the principal ideal of $\Z[\zeta]$ generated by $p$. By the Correspondence Theorem, the ideals of $\Z[\zeta]$ containing $p\Z[\zeta]$ are in bijection with the ideals of $\quotient{\Z[\zeta]}{p\Z[\zeta]}$. Since the latter object is finite, it contains finitely many ideals, meaning that only finitely many of the ideals of $\Z[\zeta]$ contain $p\Z[\zeta]$. We can then look at the maximal (proper) ideal amongst these, which is a maximal ideal of $\Z[\zeta]$. Denote it by $P$. Indeed, we can show that $P \cap \Z = p\Z$.

\begin{theorem}
    Let $g \in G$, and let $y$ be the $p'$-part of $g$ for some $p$ prime. For all characters $\chi$ of $G$, $\chi(g) - \chi(y)$ lies in the maximal ideal $P$ defined above.
\end{theorem}
\begin{proof}
    Let $m = \ord{g} = u p^v$. Let $a,b \in \Z$ be such that $au + b p^v = 1$ (as in the proof of Lemma \ref{Ch2:Lem:p_and_p'_parts}). We have that $y = g^{bp^v}$. We know that $\chi(g), \chi(y) \in \Z[\zeta]$.

    Now, let $w$ be an $m$th root of unity. Since $m \vert n$ (by Lagrange's Theorem), we know that $w \in \Z[\zeta]$. We have that $w = w^{au + bp^v}$. Raising both sides to the $p^v$th power, we get that $w^{p^v} = w^{aup^v} w^{bp^{2v}}$. Since $w^{up^v} = w^m = 1$, we have that $w^{p^v} = w^{bp^{2v}}$.

    Now, consider the binomial expansion of $\parenth{w - w^{bp^v}}^{p^v}$. Some blah blah, use fact that $P$ is a prime ideal (since it is maximal). % FINISH
\end{proof}

This theorem has several important consequences.

\begin{corollary}
    Let $g \in G$ and $y$ be the $p'$-part of $g$ for some prime $p$. Let $\chi$ be a character of $G$.
    \begin{enumerate}[label = \normalfont \arabic*., noitemsep]
        \item If $\chi(g), \chi(y) \in \Z$, then $\chi(g) \equiv \chi(y) \pmod{p}$.
        \item If $\ord{g} = p^k$ for some $k \in \N$, then $\chi(g) \equiv \chi(1) \pmod{p}$.
    \end{enumerate}
\end{corollary}

% Move the following to the end of the bit on orthogonality
\begin{lemma}
    Let $g, h \in G$. Then, $g$ is conjugate to $h$ iff for all characters $\chi$ of $G$, $\chi(g) = \chi(h)$.
\end{lemma}
\begin{proof}
    \hfill
    \begin{description}
        \item[$\parenth{\implies}$] Seen Before
        \item[$\parenth{\impliedby}$] As we know, $\Irr{G}$ is an orthonormal basis for $\FcGC$. Let $f$ be the indicator function on $g^G$. Then, $f(g) = 1 \implies f(h) = 1 \implies h \in g^G$.
    \end{description}
\end{proof}
\begin{corollary}
    Let $g \in G$. Then, $g$ is conjugate to $g^{-1}$ iff for all characters $\chi$ of $G$, $\chi(g) \in \R$.
\end{corollary}

\begin{boxexample}[Constructing the Character Table of $G = \PSL{2, 7}$]
    First, note that $\abs{G} = 168$ and that $G$ has $6$ conjugacy classes (and hence, $6$ irreducible representations over $\C$). Denote the conjugacy classes by $C_1, \ldots, C_6$, with respective representatives $g_1, \ldots, g_6$. As per convention, $g_1 = 1$. \\

    We begin filling the table as follows.
    \begin{table}[H]
        \centering
        \begin{tabular}{|c|cccccc|}
            $C_G(g_i)$ & $168$ & $8$ & $4$ & $3$ & $7$ & $7$ \\
            $\ord{g_i}$ & $1$ & $2$ & $4$ & $3$ & $7$ & $7$ \\
            $g_i$ & $g_1$ & $g_2$ & $g_3$ & $g_4$ & $g_5$ & $g_6$ \\
            \hline
            $\chi_1$ & $1$ & $1$ & $1$ & $1$ & $1$ & $1$ \\
            $\chi_2$ & & & & & & \\
            $\chi_3$ & & & & & & \\
            $\chi_4$ & & & & & & \\
            $\chi_5$ & & & & & & \\
            $\chi_6$ & & & & & & \\
        \end{tabular}
    \end{table}
    We can then fill in the first column by applying Corollary \ref{Ch2:Cor:Reg_Rep_Col_1_Grp_Size}.
    \begin{table}[H]
        \centering
        \begin{tabular}{|c|cccccc|}
            $C_G(g_i)$ & $168$ & $8$ & $4$ & $3$ & $7$ & $7$ \\
            $\ord{g_i}$ & $1$ & $2$ & $4$ & $3$ & $7$ & $7$ \\
            $g_i$ & $g_1$ & $g_2$ & $g_3$ & $g_4$ & $g_5$ & $g_6$ \\
            \hline
            $\chi_1$ & $1$ & $1$ & $1$ & $1$ & $1$ & $1$ \\
            $\chi_2$ & $7$ & & & & & \\
            $\chi_3$ & $8$ & & & & & \\
            $\chi_4$ & $3$ & & & & & \\
            $\chi_5$ & $3$ & & & & & \\
            $\chi_6$ & $6$ & & & & & \\
        \end{tabular}
    \end{table}
    Observe that since the orders are unique for the first four columns, they contain \underline{all} elements of their respective orders. We can then fill in the third column by applying the Column Orthogonality Relation and the first Lemma from today. % ADD REFERENCE! %% MOVE FILLED OUT COLUMN 1 TO THE RIGHT! IT SHOULD BE THE THIRD ONE BUT WE ACCIDENTALLY FILLED IT IN AS THE SECOND.
    \begin{table}[H]
        \centering
        \begin{tabular}{|c|cccccc|}
            $C_G(g_i)$ & $168$ & $8$ & $4$ & $3$ & $7$ & $7$ \\
            $\ord{g_i}$ & $1$ & $2$ & $4$ & $3$ & $7$ & $7$ \\
            $g_i$ & $g_1$ & $g_2$ & $g_3$ & $g_4$ & $g_5$ & $g_6$ \\
            \hline
            $\chi_1$ & $1$ & $1$ & $1$ & $1$ & $1$ & $1$ \\
            $\chi_2$ & $7$ & & $-1$ & & & \\
            $\chi_3$ & $8$ & & $0$ & & & \\
            $\chi_4$ & $3$ & & $1$ & & & \\
            $\chi_5$ & $3$ & & $1$ & & & \\
            $\chi_6$ & $6$ & & $0$ & & & \\
        \end{tabular}
    \end{table}
    We can then fill in the fourth column using the second point of Corollary 2.3.11. % ADD REFERENCE!
    \begin{table}[H]
        \centering
        \begin{tabular}{|c|cccccc|}
            $C_G(g_i)$ & $168$ & $8$ & $4$ & $3$ & $7$ & $7$ \\
            $\ord{g_i}$ & $1$ & $2$ & $4$ & $3$ & $7$ & $7$ \\
            $g_i$ & $g_1$ & $g_2$ & $g_3$ & $g_4$ & $g_5$ & $g_6$ \\
            \hline
            $\chi_1$ & $1$ & $1$ & $1$ & $1$ & $1$ & $1$ \\
            $\chi_2$ & $7$ & & $-1$ & $1$ & & \\
            $\chi_3$ & $8$ & & $0$ & $-1$ & & \\
            $\chi_4$ & $3$ & & $1$ & $0$ & & \\
            $\chi_5$ & $3$ & & $1$ & $0$ & & \\
            $\chi_6$ & $6$ & & $0$ & $0$ & & \\
        \end{tabular}
    \end{table}
    We now have four possibilities for the second column. % CHECK PHONE. SECOND COL WAS \pm EVERYWHERE, AND 2 AND 0 COULD HAVE BOTH BEEN SWAPPED.
    \begin{table}[H]
        \centering
        \begin{subtable}{0.45\linewidth}
            \begin{tabular}{|c|cccccc|}
                $C_G(g_i)$ & $168$ & $8$ & $4$ & $3$ & $7$ & $7$ \\
                $\ord{g_i}$ & $1$ & $2$ & $4$ & $3$ & $7$ & $7$ \\
                $g_i$ & $g_1$ & $g_2$ & $g_3$ & $g_4$ & $g_5$ & $g_6$ \\
                \hline
                $\chi_1$ & $1$ & $1$ & $1$ & $1$ & $1$ & $1$ \\
                $\chi_2$ & $7$ & & $-1$ & $1$ & & \\
                $\chi_3$ & $8$ & & $0$ & $-1$ & & \\
                $\chi_4$ & $3$ & & $1$ & $0$ & & \\
                $\chi_5$ & $3$ & & $1$ & $0$ & & \\
                $\chi_6$ & $6$ & & $0$ & $0$ & & \\
            \end{tabular}
        \end{subtable}
        \hspace{0.05\linewidth}
        \begin{subtable}{0.45\linewidth}
            \begin{tabular}{|c|cccccc|}
                $C_G(g_i)$ & $168$ & $8$ & $4$ & $3$ & $7$ & $7$ \\
                $\ord{g_i}$ & $1$ & $2$ & $4$ & $3$ & $7$ & $7$ \\
                $g_i$ & $g_1$ & $g_2$ & $g_3$ & $g_4$ & $g_5$ & $g_6$ \\
                \hline
                $\chi_1$ & $1$ & $1$ & $1$ & $1$ & $1$ & $1$ \\
                $\chi_2$ & $7$ & & $-1$ & $1$ & & \\
                $\chi_3$ & $8$ & & $0$ & $-1$ & & \\
                $\chi_4$ & $3$ & & $1$ & $0$ & & \\
                $\chi_5$ & $3$ & & $1$ & $0$ & & \\
                $\chi_6$ & $6$ & & $0$ & $0$ & & \\
            \end{tabular}
        \end{subtable}
    \end{table}
    For the last two columns of the table, we use row orthogonality, in combination with the fact that $g_5 = g_6\inv$.
    \begin{table}[H]
        \centering
        \begin{tabular}{|c|cccccc|}
            $C_G(g_i)$ & $168$ & $8$ & $4$ & $3$ & $7$ & $7$ \\
            $\ord{g_i}$ & $1$ & $2$ & $4$ & $3$ & $7$ & $7$ \\
            $g_i$ & $g_1$ & $g_2$ & $g_3$ & $g_4$ & $g_5$ & $g_6$ \\
            \hline
            $\chi_1$ & $1$ & $1$ & $1$ & $1$ & $1$ & $1$ \\
            $\chi_2$ & $7$ & & $1$ & $-1$ & $0$ & $0$ \\
            $\chi_3$ & $8$ & & $-1$ & $0$ & $1$ & $1$ \\
            $\chi_4$ & & & $0$ & $1$ & $z$ & $\bar{z}$ \\
            $\chi_5$ & $3$ & & $0$ & $1$ & $\bar{z}$ & $z$ \\
            $\chi_6$ & $6$ & & $0$ & $0$ & $t$ & $\bar{t}$ \\
        \end{tabular}
    \end{table}
    % CHECK PHONE FOR REST!

    FINAL COMMENTS. Also use the results on $p$ groups, since $G$ is a $7$-group.
\end{boxexample}